Carbonate chemistry at Ny-Ålesund: AWIPEV-CO~2~ Project

### 2020-05-26 awipev-CO2.Rmd
Added ESSD boxplot figure

```{r save all, echo=FALSE, message= FALSE, warning=FALSE}
# d_hour is the real data set that is read at the beginning of the script
# d_all comprises additional columns calculated in this script
d_all$at <- ifelse(is.nan(d_all$at), NA, d_all$at) # replace all NaN by NA
d_all$ct <- ifelse(is.nan(d_all$at), NA, d_all$ct) # replace all NaN by NA
saveRDS(file= paste0(path, "fb_data/d_all.rds"), d_all)
```

@@

```{r figures for ESSD paper, echo=FALSE, message= FALSE, warning=FALSE}
Mytheme <- function(size_labs = 6, face_font="plain", ...) {
  theme_bw() +
  theme(
    # axis.text.x = element_text(face=face_font, size=size_labs, color="black"),
    #     axis.title.x = element_text(face="plain", size=size_labs, margin=margin(0,0,0,0,"pt")),
    #     axis.text.y = element_text(face=face_font, color="black", size=size_labs),
    axis.text.x = element_blank(),
    #     axis.title.y = element_text(face="plain", size=size_labs),
    #     axis.ticks.x = element_line(size=0.1),
    #     axis.ticks.y = element_line(size=0.1),
    #     axis.ticks.length = unit(1.1, "mm"),
    #     panel.grid.major = element_line(size = 0.2, color="black", linetype="dotted"),
    #     aspect.ratio = 1 / 3,
        plot.margin = margin(t = 0, r = 0.3, b = 0, l = 0.3, unit = "cm")
  )
}

xlim <- c(as.POSIXct("2015-07-01 00:00:00"), as.POSIXct("2020-07-01 00:00:00"))
# calculate omeaga arag
z <- d_all %>% 
  dplyr::filter(!(is.na(pco2_corr) | is.na(at_calc) | is.na(sal_mix) |is.na(temp_insitu_11m) | is.na(pressure_insitu_ctd)))
oa_tbf <- as_tibble(
  carb(24, z$pco2_corr, z$at_calc*1e-6, S = z$sal_mix, T = z$temp_insitu_11m, 
                          P = z$pressure_insitu_ctd/10, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", 
                          ks = "d", pHscale = "T", b = "u74")
) %>% 
  dplyr::mutate(datetime = z$datetime,
                oa = OmegaAragonite) %>% 
dplyr::select(datetime, oa)
# select time series data
d <- as_tibble(d_all) %>%
  dplyr::select(datetime, pressure_insitu_ctd, temp_insitu_11m, sal_mix, pco2_corr, pHint_tot_sf, 
                ph_dur_corr, at_calc)
d <- left_join(d, oa_tbf, by = 'datetime')

# time series temp
ts_temp <- d %>%
  dplyr::filter(!is.na(temp_insitu_11m)) %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = temp_insitu_11m), col = "blue", size = 0.15) +
  labs(title = "", x = "", y = "Temp. 11 m (°C)") +
  Mytheme(size_labs = 8)
print(ts_temp) 
# time series sal
ts_sal <- d %>%
  dplyr::filter(!is.na(sal_mix)) %>% 
ggplot() +
  geom_point(aes(x = datetime, y = sal_mix), col = "blue", size = 0.15) + 
  labs(title = "", x = "", y = "Salinity") +
  Mytheme(size_labs = 8)
print(ts_sal) 
# time series pco2
ts_pco2 <- d %>%
  dplyr::filter(!is.na(pco2_corr)) %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = pco2_corr), col = "blue", size = 0.15) +
  scale_x_datetime(limits = xlim) +
  labs(title = "", x = "", y = expression(paste("pC", O[2], " (", mu, "atm)"))) +
  Mytheme(size_labs = 8)
print(ts_pco2) 
# time series pH
ts_ph <- d %>%
  dplyr::select(datetime, pHint_tot_sf, ph_dur_corr) %>%
  dplyr::filter(!is.na(pHint_tot_sf)) %>%
  pivot_longer(-datetime, names_to = "pH", values_to = "value") %>%
  ggplot() +
  geom_point(aes(x = datetime, y = value, colour = pH), size = 0.15) +
  scale_x_datetime(limits = xlim) +
  labs(title = "",
       x = "",
       y = expression(paste("p", H[T]))) +
  # scale_color_manual(
  #   values = c("blue", "red"),
  #   name = "",
  #   breaks = c("pHint_tot_sf", "ph_dur_corr"),
  #   labels = c("SeaFET", "Durafet")
  # ) +
  annotate(
    geom = "text",
    x = as.POSIXct("2018-06-01"),
    y = 8.35,
    label = "seaFET",
    color = "blue", size = 2
  ) +
  annotate(
    geom = "text",
    x = as.POSIXct("2018-06-01"),
    y = 8.25,
    label = "Durafet",
    color = "red", size = 2
  ) +
  Mytheme(size_labs = 8) +
  theme(legend.position = "none")
print(ts_ph) 
# time series AT
ts_at <- d %>%
  dplyr::filter(!is.na(at_calc)) %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = at_calc), col = "blue", size = 0.15) +
  scale_x_datetime(limits = xlim) +
  labs(title = "", x = "Time", y = expression(paste(italic(A)[T]," ","(",mu, mol.kg^-1,")"))) +
  Mytheme(size_labs = 8) +
  theme(axis.title.x = element_text(face="plain", size=8))
print(ts_at) 
# time series Omega
ts_oa <- d %>%
  dplyr::filter(!is.na(oa)) %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = oa), col = "blue", size = 0.15) +
  scale_x_datetime(limits = xlim) +
  labs(title = "", x = "Time", y = "Omega aragonite") +
  Mytheme(size_labs = 8) +
  theme(axis.title.x = element_text(face="plain", size=8))
print(ts_oa) 
#assemble time-series figure
cp <- cowplot::plot_grid(ts_sal, ts_temp, ts_pco2, ts_ph, ts_at, ts_oa,
                   align = "v",
                   #labels = "AUTO",
                   ncol = 1
                   #width = 18,
                   #units = "cm"
                   )
print(cp)
cowplot::ggsave2(filename = "figures/ts_gg.png", plot = cp, height = 20, units = "cm")

# ts_sal / ts_temp / ts_at
# grid.arrange(ts_sal, ts_temp, ts_pco2, ts_ph, ts_at, ncol = 1)
# cowplot::save_plot(filename = "figures/ts_sp.pdf", plot = cp)
# multiplot(ts_sal, ts_temp, ts_pco2, ts_ph, ts_at, cols = 1)
```

<!-- ```{r Klara Wolf} -->
<!-- wolf <- d_all %>% -->
<!--   dplyr::select(datetime, PCO2_Corr_filtered, phcalc, pH_calc_norm4) %>% -->
<!--   dplyr::filter(datetime >= "2016-03-01" & datetime <= "2016-09-30") -->
<!-- write_csv(path = "wolf.csv", wolf, col_names = TRUE) -->
<!-- ``` -->

>>>>

```{r save all, echo=FALSE, message= FALSE, warning=FALSE}
# d_hour is the real data set that is read at the beginning of the script
# d_all comprises additional columns calculated in this script
d_all$at <- ifelse(is.nan(d_all$at), NA, d_all$at) # replace all NaN by NA
d_all$ct <- ifelse(is.nan(d_all$at), NA, d_all$ct) # replace all NaN by NA
  dplyr::mutate(month = as.factor(month(x = datetime, label = TRUE))) %>%

saveRDS(file= paste0(path, "fb_data/d_all.rds"), d_all)
```

@@

```{r ESSD time series, echo=FALSE, message= FALSE, warning=FALSE}
mytheme <- function(size_labs = 6, face_font="plain", ...) {
  theme_bw() +
  theme(axis.text.x = element_text(face=face_font, size=size_labs, color="black"),
        axis.title.x = element_text(face=face_font, size=size_labs, margin=margin(0,0,0,0,"pt")),
        axis.text.y = element_text(face=face_font, color="black", size=size_labs),
        axis.title.y = element_text(face=face_font, size=size_labs),
        axis.ticks.x = element_line(size=0.1),
        axis.ticks.y = element_line(size=0.1),
        axis.ticks.length = unit(1.1, "mm"),
        panel.grid.major = element_line(size = 0.25, color="black", linetype="dotted"),
        aspect.ratio = 1 / 3,
        plot.margin = margin(t = 0, r = 1, b = 0, l = 0, unit = "lines")
  )
}

xlim <- c(as.POSIXct("2015-07-01 00:00:00"), as.POSIXct("2020-07-01 00:00:00"))
# calculate omeaga arag
z <- d_all %>% 
  dplyr::filter(!(is.na(pco2_corr) | is.na(at_calc) | is.na(sal_mix) |is.na(temp_insitu_11m) | is.na(pressure_insitu_ctd)))
oa_tbf <- as_tibble(
  carb(24, z$pco2_corr, z$at_calc*1e-6, S = z$sal_mix, T = z$temp_insitu_11m, 
                          P = z$pressure_insitu_ctd/10, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", 
                          ks = "d", pHscale = "T", b = "u74")
) %>% 
  dplyr::mutate(datetime = z$datetime,
                oa = OmegaAragonite) %>% 
dplyr::select(datetime, oa)
# select time series data and add Omega
d <- as_tibble(d_all) %>%
  dplyr::select(datetime, pressure_insitu_ctd, temp_insitu_11m, sal_mix, pco2_corr, pHint_tot_sf, 
                ph_dur_corr, at_calc)
d <- left_join(d, oa_tbf, by = 'datetime')

# time series temp
ts_temp <- d %>%
  dplyr::filter(!is.na(temp_insitu_11m)) %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = temp_insitu_11m), col = "blue", size = 0.15) +
  labs(title = "", x = "", y = "Temp. 11 m (°C)") +
  mytheme(size_labs = 8)
print(ts_temp) 
# time series sal
ts_sal <- d %>%
  dplyr::filter(!is.na(sal_mix)) %>% 
ggplot() +
  geom_point(aes(x = datetime, y = sal_mix), col = "blue", size = 0.15) + 
  labs(title = "", x = "", y = "Salinity") +
  mytheme(size_labs = 8)
print(ts_sal) 
# time series pco2
ts_pco2 <- d %>%
  dplyr::filter(!is.na(pco2_corr)) %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = pco2_corr), col = "blue", size = 0.15) +
  scale_x_datetime(limits = xlim) +
  labs(title = "", x = "", y = expression(paste("pC", O[2], " (", mu, "atm)"))) +
  mytheme(size_labs = 8)
print(ts_pco2) 
# time series pH
ts_ph <- d %>%
  dplyr::select(datetime, pHint_tot_sf, ph_dur_corr) %>%
  dplyr::filter(!is.na(pHint_tot_sf)) %>%
  pivot_longer(-datetime, names_to = "pH", values_to = "value") %>%
  ggplot() +
  geom_point(aes(x = datetime, y = value, colour = pH), size = 0.15) +
  scale_x_datetime(limits = xlim) +
  labs(title = "",
       x = "",
       y = expression(paste("p", H[T]))) +
  # scale_color_manual(
  #   values = c("blue", "red"),
  #   name = "",
  #   breaks = c("pHint_tot_sf", "ph_dur_corr"),
  #   labels = c("SeaFET", "Durafet")
  # ) +
  annotate(
    geom = "text",
    x = as.POSIXct("2018-06-01"),
    y = 8.35,
    label = "seaFET",
    color = "blue", size = 2
  ) +
  annotate(
    geom = "text",
    x = as.POSIXct("2018-06-01"),
    y = 8.25,
    label = "Durafet",
    color = "red", size = 2
  ) +
  mytheme(size_labs = 8) +
  theme(legend.position = "none")
print(ts_ph) 
# time series AT
ts_at <- d %>%
  dplyr::filter(!is.na(at_calc)) %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = at_calc), col = "blue", size = 0.15) +
  scale_x_datetime(limits = xlim) +
  labs(title = "", x = "Time", y = expression(paste(italic(A)[T]," ","(",mu, mol.kg^-1,")"))) +
  mytheme(size_labs = 8) +
  theme(axis.title.x = element_text(face="plain", size=8))
print(ts_at) 
# time series Omega
ts_oa <- d %>%
  dplyr::filter(!is.na(oa)) %>% 
  ggplot() +
  geom_point(aes(x = datetime, y = oa), col = "blue", size = 0.15) +
  scale_x_datetime(limits = xlim) +
  labs(title = "", x = "Time", y = expression(paste(Omega[a]))) +
  Mytheme(size_labs = 8) +
  theme(axis.title.x = element_text(face="plain", size=8))
print(ts_oa) 
#assemble time-series figure
cp <- cowplot::plot_grid(ts_sal, ts_temp, ts_pco2, ts_ph, ts_at, ts_oa,
                   align = "v",
                   #labels = "AUTO",
                   ncol = 1
                   #width = 18,
                   #units = "cm"
                   )
print(cp)
cowplot::ggsave2(filename = "figures/ts_gg.png", plot = cp, height = 20, units = "cm")

# ts_sal / ts_temp / ts_at
# grid.arrange(ts_sal, ts_temp, ts_pco2, ts_ph, ts_at, ncol = 1)
# cowplot::save_plot(filename = "figures/ts_sp.pdf", plot = cp)
# multiplot(ts_sal, ts_temp, ts_pco2, ts_ph, ts_at, cols = 1)
```
```{r ESSD boxplots, echo=FALSE, message= FALSE, warning=FALSE}
mytheme_bp <- function(size_labs = 9, face_font="plain", ...) {
  mytheme() +
    theme_bw() +
  theme(aspect.ratio = 1/2,
        plot.margin = margin(t = -2, r = 0.2, b = -2, l = 0, unit = "cm"),
        axis.text.x=element_blank())
  }
d_long <- d %>% 
  dplyr::mutate(month = as.factor(month(x = datetime, label = TRUE))) %>% 
  tidyr::pivot_longer(-c(datetime, month), names_to = "variable", values_to = "value")

sal_mix_bp <- ggplot(filter(d_long, variable=="sal_mix"), aes(x = month, y = value, group=month)) +
  geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE) +
  geom_boxplot(notch=TRUE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  scale_x_discrete(breaks=c("Jan", "", "Mar", "", "May", "", "Jul", "", "Sep", "", "Nov", "")) +
  labs(title=NULL,x=NULL,y = "Salinity") +
  mytheme_bp()
temp_11m_bp <- ggplot(filter(d_long, variable=="temp_insitu_11m"), aes(x = month, y = value, group=month)) +
  geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE) +
  geom_boxplot(notch=TRUE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  scale_x_discrete(breaks=c("Jan", "", "Mar", "", "May", "", "Jul", "", "Sep", "", "Nov", "")) +
  labs(title=NULL,x=NULL,y = "Temp. 11 m (°C)") +
  mytheme_bp()
pco2_bp <- ggplot(filter(d_long, variable=="pco2_corr"), aes(x = month, y = value, group=month)) +
  geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE) +
  geom_boxplot(notch=TRUE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  scale_x_discrete(breaks=c("Jan", "", "Mar", "", "May", "", "Jul", "", "Sep", "", "Nov", "")) +
  labs(title=NULL,x=NULL,y = expression(paste("pC", O[2], " (", mu, "atm)"))) +
  mytheme_bp()
ph_bp <- ggplot(filter(d_long, variable=="pHint_tot_sf"), aes(x = month, y = value, group=month)) +
  geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE) +
  geom_boxplot(notch=TRUE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  scale_x_discrete(breaks=c("Jan", "", "Mar", "", "May", "", "Jul", "", "Sep", "", "Nov", "")) +
  labs(title=NULL,x=NULL,y = expression(paste("p", H[T]))) +
  mytheme_bp()
at_bp <- ggplot(filter(d_long, variable=="at_calc"), aes(x = month, y = value, group=month)) +
  geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE) +
  geom_boxplot(notch=TRUE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  scale_x_discrete(breaks=c("Jan", "", "Mar", "", "May", "", "Jul", "", "Sep", "", "Nov", "")) +
  labs(title=NULL,x=NULL,y = expression(paste(italic(A)[T]," ","(",mu, mol.kg^-1,")"))) +
  mytheme_bp() +
  theme(axis.text.x = element_text(vjust = 1))
oa_bp <- ggplot(filter(d_long, variable=="oa"), aes(x = month, y = value, group=month)) +
  geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE) +
  geom_boxplot(notch=TRUE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  scale_x_discrete(breaks=c("Jan", "", "Mar", "", "May", "", "Jul", "", "Sep", "", "Nov", "")) +
  labs(title=NULL,x=NULL,y = expression(paste(Omega[a]))) +
  mytheme_bp() +
  theme(axis.text.x = element_text(vjust = 1))

g <- cowplot::plot_grid(sal_mix_bp, temp_11m_bp, pco2_bp, ph_bp, at_bp, oa_bp, ncol=2,
               align="v")
g
#g <- arrangeGrob(T1vp, S1vp, AT1vp, CT1vp, pH1vp, pCO21vp, Oa1vp, ncol=1)
#cowplot::
ggsave(file="figures/boxplots.png", g,  width = 18, height = 14, units = "cm")
#g <- cowplot::plot_grid(T1vp, S1vp, AT1vp, CT1vp, pH1vp, pCO21vp, Oa1vp, ncol=2)
#ggsave(file="figures/fig_paper_monthly_means_1m.png", g, width = 18, height = 14, units = "cm")
```
